Railway: Adaptive Disk Storage for Interaction Graphs

### 2014-09-17 docs/edbt2015/text/optimal.tex
started the optimal railway design section

\section{Optimal Railway Design}\label{sec:optimal}  
The optimal railway design concerns the partitioning of the disk blocks into
sub-blocks such that the query I/O is minimized, while the storage overhead
induced is kept below a specified threshold. This optimization is guided by
the query workload, as observed by a particular disk block at hand.

The partitioning of disk blocks into sub-blocks can be performed in two ways:
\emph{non-overlapping} and \emph{overlapping}. In the non-overlapping case,
each sub-block contains a non-overlapping subset of the edge attributes. In
contrast, in the overlapping case, the attributes contained within sub-blocks
can overlap. In both cases, the complete graph structure for the block is
replicated in the sub-blocks, which is the main cause of increase in the query
I/O.

In both overlapping and non-overlapping partitioning, we trade increased
storage  overhead for reduced space. In the overlapping case, the increase in
the storage overhead is higher, as the space used for some of the attributes
are replicated  as well, in addition to the graph structure. On the other
hand, enabling overlapping attributes is expected to improve the query I/O.
While the non-overlapping partitioning scenario is a special case of the more
general case of overlapping partitioning, more specialized algorithms can be
used to solve the non-overlapping partitioning problem.

In the rest of this section, we first introduce some basic notation, and then
formulate the overall optimization problem. The formulation of the query I/O
cost and  storage overhead are presented next to complete the definition of
the optimal railway design problem.

HERE

>>>>

\section{Optimal Railway Design}\label{sec:optimal}  
The optimal railway design concerns the partitioning of the disk blocks into
sub-blocks such that the query I/O is minimized, while the storage overhead
induced is kept below a desired threshold. This optimization is guided by the
query workload, as observed by a disk block within a particular time range.

The partitioning of disk blocks into sub-blocks can be \emph{non-overlapping}
or \emph{overlapping}. In the non-overlapping case, each sub-block contains a
non-overlapping subset of the edge attributes. In the overlapping case, the
subset of attributes contained within sub-blocks can overlap. In both cases,
the complete graph structure for the block is replicated in the sub-blocks,
which is the main cause of the increased storage overhead.

In both overlapping and non-overlapping partitioning, we trade increased
storage overhead for reduced query I/O cost. In the overlapping case, the
increase in the storage overhead is higher, as some of the attributes are
replicated, in addition to the graph structure. On the other hand, enabling
overlapping attributes is expected to reduce the query I/O (in the extreme
case, there could be one sub-block per query). While the non-overlapping
partitioning scenario is a special case of the  overlapping one, specialized
algorithms can be used to solve the former problem.

In the rest of this section, we first introduce basic notation and then
formulate the overall optimization problem. The modeling of the query I/O and
storage overhead are presented next, which complete the formalization of the
optimal railway design problem.
